---
title: "phlbbycatcher"
output: rmarkdown::html_vignette
toc:
  depth: 1
vignette: >
  %\VignetteIndexEntry{phlbbycatcher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(phlbbycatcher)
```
# Introduction
This is a quick start guide and, as such, is only intended to give the minimal instructions necessary to complete the annual reporting. For more in-depth discussion of annual reporting, please see: "Creating Annual P. halibut Bycatch Estimates with **phlbbycatcher**".

The **phlbbycatcher** package aims to simplify the analysis and reporting of P. halibut bycatch mortality to the PFMC and other stakeholders.  Specifically, the package breaks the process down into the following steps:

 1. Stratify the data.  Not all fisheries are stratified in the same manner.
 2. Create observer coverage tables, stratified in the same manner as step 1 above.  This step also tests for and masks confidential strata.
 3. Estimate P. halibut bycatch mortality.
 4. Format output, masking confidential data when necessary.
 
 This document introduces you to **phlbbycatcher**'s basic functions and shows how to apply them to U.S. west coast fisheries data from a variety of sources. The organization and underlying functions incorporates the design philosophy of the [`tidyverse`](https://www.tidyverse.org/) to the greatest extent possible.

# 1. Set up Network Folders
1. Create new folders for `data.year` in 
    * `/nwctantalus.nmfs.local/observer/Input/PHLB/<<data.year>>/`
    * `/nwctantalus.nmfs.local/observer/Input/PHLB/<<data.year>>/analyst_data/`
    * `/nwctantalus.nmfs.local/observer/Input/PHLB/<<data.year>>/Inseason/`
    * `/nwctantalus.nmfs.local/observer/Output/Pacific Halibut <<data.year>>/`
2. Add data from requests (see below) to:
    * `/nwctantalus.nmfs.local/observer/Input/PHLB/<<data.year>>/`

## File Paths
File paths should be "auto-magically" created and should not need to be set by the analyst, *except* for `tant_drive` (see [4. Update `000_Setup.Rmd`]{#Setup} below).  The file paths for the annual report are determined by the environment in which R is being run (local machine vs. `tantalus` server) which is identified in `000_Setup.Rmd` (or `Inseason_Setup.Rmd`). Drives should be created appropriately for both `tantalus` as well as your local machine. For more details, see "Creating Annual P. halibut Bycatch Estimates with **phlbbycatcher**".
    
# 2. Obtain Required Data Inputs
Ensure you have the following inputs:    

1. FOS data, obtained  from the `load_data_modified()` function:    
 * `OBOrig_Pre`,`OBOrig_Proc`    
 * `SPC`    
 * `FTOrig_Pre`, `FTOrig_Proc`    
 * `ASOrig_Proc`    
 * `EMOrig_Pre`, `EMOrig_Proc`    
 
 2. WCGOP PHLB Biosamples data from FOS Data Team (currently James Fellows).    
 3. A-SHOP PHLB Length data from A-SHOP (currently V. Tuttle).    
 
 4. EM PHLB length and time-on-deck data from from PSMFC.  Typically, this is included in an annual request from the FOS Analyst team, and can usually be found here: 
 
 <!-- FIX THIS -->
 `/FRAM/Observer/observer/..........`    
 
 Place all data from above in:    
 `/nwctantalus.nmfs.local/observer/Input/PHLB/<<data.year>>/`    

 `Annual_Data.R` reads the raw data from the above location and then repackages it as `.Rda` file and then saves it in:    
 `/nwctantalus.nmfs.local/observer/Output/Pacific Halibut <<data.year>>/`

# 3. Update Dates
Update the following dates in `scripts/Dates.R`:    

    * `aslengthdate`    
    * `cscovdte`    
    * `emcovdte`    
    * `ncscovdte`    
    
The `aslengthdate` is the date in the filename of the A-SHOP P. halibut lengths.  The other names are dates in the filenames of the coverage products produced by the FOS analysts and found here:

 <!-- FIX THIS -->
 `/nwctantalus.nmfs.loca/observer/Output/Coverage/2002-<<data.year>>/`    
 

# 4. Update `000_Setup.Rmd` {#Setup}
* define the logicals that control how code is run    
* define location and time specific constants (e.g, drives and data.year)    
    - `tant_drive` - Define the drive where `tantalus` is located, e.g.: `tant_drive <- "V:/"`. Note that, "V:/" is J. Jannot's drive letter for `tantalus` - *change this letter to match your own drive letter* for `tantalus`.    
    
# 5. Knit `0000_PHLB_Report.Rmd`
This script is the controller script which calls and implements all other `*.Rmd` and `*.R` scripts.

# 6. Troubleshoot
If non-obvious trouble arises, check `000_Setup.Rmd` for errors.

***
# After Knitting
## 1. Check Report
Ensure report text, tables, figures, references and appendices are correct.

## 2. Check Supplementary Tables
Ensure the MSExcel Workbook of Supplementary tables are correct, with captions, headers, and correct sheet numbering corresponding to table numbers.

***

# Reviewing & Submitting
1. Send for reviews by co-authors.
2. Send for review by IPHC.
3. Submit, annually in August, to PFMC September Briefing Book. This can occur before recieving IPHC review.
4. Submit to RPTS, with IPHC review as the Technical Review.  Add Craig Russell as Technical Reviewer.


***
```{r outro, include=FALSE}
pkg <- "knitr"
pkg.version <- installed.packages()[pkg, 'Version']

pkg.info <- paste("`knitr` version", pkg.version, sep=" ")
Rver <- paste(R.Version()$version.string, " x", strsplit(R.Version()$arch, "_")[[1]][2], sep = "")
os <- paste(as.character(Sys.info()[1:2]), collapse = " ")
```
This document was processed by `r pkg.info` using `r Rver` on a `r os` machine. 
